---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Freedom House Datasets in R

<!-- badges: start -->
<!-- badges: end -->

The goal of freedomhouse is to ease the usage of the Freedom in the World
dataset from Freedom House in R. The Freedom in the World dataset is updated
annually and is originally available for download in Excel format.

Freedom House is best known for political advocacy surrounding issues of 
Democracy, Political Freedom, and Human Rights. Each of these issues trascends
political colours. The Freedom in the World dataset is a comprehensive and
widely used measure of political freedom. It is used by academics, journalists,
and policy makers alike.

My added value is to presents all the tables in a really simple to use format
and to make all the texts with the justifications for each sub-item scores,
that you find in around 1,000 links of the form 
https://freedomhouse.org/country/canada/freedom-world/2023, available in a
single tidy table. This is useful for text mining and sentiment analysis.

Another addition of mine was is to add translations, such as the side package
`casadelalibertad`.

## Installation

You can install the development version of freedom from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("pachadotdev/freedom")
```

To install a translation, such as the Spanish translation, you can use:

```r
remotes::install_github("pachadotdev/casadelalibertad", subdir = "translations/es")
```

## Example

This is a basic example which shows you how to join three tables

```{r example, warning=FALSE, message=FALSE}
library(dplyr)
library(freedomhouse)

# Search for "trade union" in the sub_item_description column
category_scores_items %>%
  filter(grepl("trade union", sub_item_description))

# Get the full description of the sub-item
category_scores_items %>%
  filter(sub_item == "E3") %>%
  pull(sub_item_description)

# Filter by sub-item code and country code for trade unions in Canada
category_scores %>%
  filter(
    sub_item == "E3",
    iso3c == "CAN"
  )  %>%
  inner_join(
    country_ratings_texts %>%
      select(year, iso3c, sub_item, detail) %>%
      filter(
        sub_item == "E3",
        iso3c == "CAN"
      ),
    by = c("year", "iso3c", "sub_item")
  ) %>%
  select(year, iso3c, sub_item, score, detail)
```

## Translations

### Spanish

## Traducciones

Las traducciones disponibles dentro de `casadelalibertad` son las siguientes:

```{r es1, echo = FALSE}
library(purrr)
library(yaml)

specs <- list.files("inst/specs")
all <- map(file.path("inst", "specs", specs), read_yaml)
sources <- map_chr(all, ~ .x$df$source)

pkgs <- tibble(
  Nombre = map_chr(all, ~ .x$df$name),
  Titulo = map_chr(all, ~ .x$help$title),
  Dataset = map_chr(strsplit(sources, "::"), ~ .x[[2]])
)

pkgs %>%
  arrange(Dataset) %>%
  knitr::kable()
```

El paquete `casadelalibertad` se carga igual que todos los paquetes de R: 

```{r es2}
library(casadelalibertad)
library(dplyr)
```

Las variables que contienen los datos van a estar disponibles inmediatamente
para su uso, pero los datos no se traducirán hasta que la variable sea "llamada"
explícitamente en el código que se escriba. En este ejemplo, el _dataset_
`puntajes_por_categoria`, que proviene de `freedomhouse::category_scores`, se 
carga en la memoria de R en el momento en que lo llamamos por primera vez en 
español:

```{r es3}
glimpse(puntajes_por_categoria)
```

Los datos traducidos quedarán cargados durante toda la sesión de R:

```{r es4}
puntajes_por_categoria %>%
  filter(pais_territorio == "Canad\u00e1")
```

## Development

(This is a note for myself)

1. Open `translations/es`.

2. Document:

```r
devtools::load_all()
writeLines(create_rd("inst/specs/category_scores.yml"),
  "man/puntajes_por_categoria.Rd")
```

3. Verify:

```r
attachment::att_amend_desc()
devtools::check()
```

4. Fix non-ASCII characters:

```bash
casadelalibertad (main) $ bash dev/fix_non_ascii.sh 
```
