---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# freedom

<!-- badges: start -->
<!-- badges: end -->

The goal of freedom is to ease the usage of the Freedom in the World dataset from Freedom House in R.

Freedom House is best known for political advocacy surrounding issues of Democracy, Political Freedom, and Human Rights.
Each of these issues trascends political colours. The Freedom in the World dataset is a comprehensive and widely used
measure of political freedom. It is used by academics, journalists, and policy makers alike.

The Freedom in the World dataset is updated annually and is originally available for download in Excel format.

## Installation

You can install the development version of freedom from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("pachadotdev/freedom")
```

## Example

This is a basic example which shows you how to join two tables

```{r example, warning=FALSE, message=FALSE}
library(dplyr)
library(freedom)

# Search for "trade union" in the sub_item_description column
category_scores_items %>%
  filter(grepl("trade union", sub_item_description))

# Get the full description of the sub-item
category_scores_items %>%
  filter(sub_item == "E3") %>%
  pull(sub_item_description)

# Filter by sub-item code and country code for trade unions in Canada
category_scores %>%
  filter(
    sub_item == "E3",
    iso3c == "CAN"
  )
```
