---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Freedom House Datasets in R

<!-- badges: start -->
<!-- badges: end -->

The goal of freedomhouse is to ease the usage of the Freedom in the World
dataset from Freedom House in R. The Freedom in the World dataset is updated
annually and is originally available for download in Excel format.

Freedom House is best known for political advocacy surrounding issues of 
Democracy, Political Freedom, and Human Rights. Each of these issues trascends
political colours. The Freedom in the World dataset is a comprehensive and
widely used measure of political freedom. It is used by academics, journalists,
and policy makers alike.

My added value is to presents all the tables in a really simple to use format
and to make all the texts with the justifications for each sub-item scores,
that you find in around 1,000 links of the form 
https://freedomhouse.org/country/canada/freedom-world/2023, available in a
single tidy table. This is useful for text mining and sentiment analysis.

Another addition of mine was is to add translations.

## Installation

You can install the development version of freedom from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("pachadotdev/freedomhouse")
```

To install a translation, such as the Spanish translation, you can use:

```r
remotes::install_github("pachadotdev/freedomhouse", subdir = "translations/es")
```

## Example

This is a basic example which shows you how to join three tables

```{r example, warning=FALSE, message=FALSE}
library(dplyr)
library(freedomhouse)

# Search for "trade union" in the sub_item_description column
country_score %>%
  filter(grepl("trade union", sub_item_description))

# Get the full description of the sub-item
country_score %>%
  filter(sub_item == "E3") %>%
  distinct(sub_item_description) %>%
  pull(sub_item_description)

# Filter by sub-item code and country code for trade unions in Canada
country_score %>%
  filter(
    sub_item == "E3",
    iso3c == "CAN"
  ) %>%
  inner_join(
    country_rating_text %>%
      select(year, iso3c, sub_item, detail) %>%
      filter(
        sub_item == "E3",
        iso3c == "CAN"
      ),
    by = c("year", "iso3c", "sub_item")
  ) %>%
  select(year, iso3c, sub_item, score, detail)
```

## Shiny

There is an example with Shiny [here](https://github.com/pachadotdev/freedomhouse/tree/main/shiny-demo).

## Translations

### Templates in Excel

The directory `dev/texts` contains the Excel files that I used for the
translations in `dev/texts/translation_es`. These can be used as templates for
translations to other languages different from Spanish.

What I did to add the translations in R was to use the `left_join` function
from the `dplyr` package to match each country/item/sub-item with the
corresponding translation.

### Spanish

El _dataset_ `puntaje_pais` es una traducción del original `country_score`:

```r
> glimpse(puntaje_pais)
Rows: 62,875
Columns: 10
$ anio                      <int> 2012, 2012, 2012, 2012, 2012, 2012, 2012, 20…
$ pais                      <chr> "Abjasia", "Abjasia", "Abjasia", "Abjasia", …
$ iso2c                     <fct> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, …
$ iso3c                     <fct> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, …
$ continente                <chr> "Asia", "Asia", "Asia", "Asia", "Asia", "Asi…
$ categoria                 <fct> A, A, A, B, B, B, B, C, C, C, D, D, D, D, E,…
$ sub_categoria             <fct> A1, A2, A3, B1, B2, B3, B4, C1, C2, C3, D1, …
$ descripcion_categoria     <chr> "Derechos políticos", "Derechos políticos", …
$ descripcion_sub_categoria <chr> "¿Fue el actual jefe de gobierno u otra auto…
$ puntaje                   <int> 3, 2, 2, 2, 2, 2, 1, 1, 1, 2, 2, 2, 1, 3, 2,…
```

```r
> puntaje_pais %>%
  filter(pais == "Canad\u00e1")
# A tibble: 300 × 10
    anio pais   iso2c iso3c continente   categoria sub_categoria
   <int> <chr>  <fct> <fct> <chr>        <fct>     <fct>        
 1  2012 Canadá CA    CAN   Las Américas A         A1           
 2  2012 Canadá CA    CAN   Las Américas A         A2           
 3  2012 Canadá CA    CAN   Las Américas A         A3           
 4  2012 Canadá CA    CAN   Las Américas B         B1           
 5  2012 Canadá CA    CAN   Las Américas B         B2           
 6  2012 Canadá CA    CAN   Las Américas B         B3           
 7  2012 Canadá CA    CAN   Las Américas B         B4           
 8  2012 Canadá CA    CAN   Las Américas C         C1           
 9  2012 Canadá CA    CAN   Las Américas C         C2           
10  2012 Canadá CA    CAN   Las Américas C         C3           
# ℹ 290 more rows
# ℹ 3 more variables: descripcion_categoria <chr>,
#   descripcion_sub_categoria <chr>, puntaje <int>
# ℹ Use `print(n = ...)` to see more rows
```

```r
> puntaje_pais %>%
  filter(pais == "Canadá")
# A tibble: 300 × 10
    anio pais   iso2c iso3c continente   categoria sub_categoria
   <int> <chr>  <fct> <fct> <chr>        <fct>     <fct>        
 1  2012 Canadá CA    CAN   Las Américas A         A1           
 2  2012 Canadá CA    CAN   Las Américas A         A2           
 3  2012 Canadá CA    CAN   Las Américas A         A3           
 4  2012 Canadá CA    CAN   Las Américas B         B1           
 5  2012 Canadá CA    CAN   Las Américas B         B2           
 6  2012 Canadá CA    CAN   Las Américas B         B3           
 7  2012 Canadá CA    CAN   Las Américas B         B4           
 8  2012 Canadá CA    CAN   Las Américas C         C1           
 9  2012 Canadá CA    CAN   Las Américas C         C2           
10  2012 Canadá CA    CAN   Las Américas C         C3           
# ℹ 290 more rows
# ℹ 3 more variables: descripcion_categoria <chr>,
#   descripcion_sub_categoria <chr>, puntaje <int>
# ℹ Use `print(n = ...)` to see more rows
```

`texto_calificacion_pais` es una traducción del original `country_rating_text`:

```r
> glimpse(texto_calificacion_pais)
Rows: 29,587
Columns: 7
$ anio          <int> 2017, 2017, 2017, 2017, 2017, 2017, 2017, 2017, 2017, 20…
$ pais          <chr> "Afganistán", "Afganistán", "Afganistán", "Afganistán", …
$ iso2c         <fct> AF, AF, AF, AF, AF, AF, AF, AF, AF, AF, AF, AF, AF, AF, …
$ iso3c         <fct> AFG, AFG, AFG, AFG, AFG, AFG, AFG, AFG, AFG, AFG, AFG, A…
$ continente    <chr> "Asia", "Asia", "Asia", "Asia", "Asia", "Asia", "Asia", …
$ sub_categoria <fct> Resumen, Desarrollos Clave, Cambio de Calificación, A1, …
$ detalle       <chr> "La constitución de Afganistán prevé un Estado unitario,…
```

`estado_calificacion_pais` es una traducción del original `country_rating_status`:

```r
> glimpse(estado_calificacion_pais)
Rows: 9,238
Columns: 9
$ anio               <int> 1972, 1973, 1974, 1975, 1976, 1977, 1978, 1979, 198…
$ pais               <chr> "Afganistán", "Afganistán", "Afganistán", "Afganist…
$ iso2c              <fct> AF, AF, AF, AF, AF, AF, AF, AF, AF, AF, AF, AF, AF,…
$ iso3c              <fct> AFG, AFG, AFG, AFG, AFG, AFG, AFG, AFG, AFG, AFG, A…
$ continente         <chr> "Asia", "Asia", "Asia", "Asia", "Asia", "Asia", "As…
$ derechos_politicos <int> 4, 7, 7, 7, 7, 6, 7, 7, 7, 7, 7, 7, 7, 7, 7, 6, 7, …
$ libertades_civiles <int> 5, 6, 6, 6, 6, 6, 7, 7, 7, 7, 7, 7, 7, 7, 7, 6, 7, …
$ estado             <chr> "Parcialmente Libre", "No Libre", "No Libre", "No L…
$ color              <fct> #a1aafc, #7454a6, #7454a6, #7454a6, #7454a6, #7454a…
```

## Development

To verify the changes made to the package, run the following commands:

```r
devtools::document()
attachment::att_amend_desc()
devtools::check()
```
